\documentclass{ck-article}

\title{Loops on the affine line}
\author{Clemens Koppensteiner}

\usepackage{math-ag}

\addbibresource{math.bib}
%\bibliography{math.bib}

\newcommand\Gpd[2]{G_{#1/#2}}

\begin{document}
\maketitle

We want to prove some things about modules over the space
\[
    Ω_0 \as1 = \pt ×_{\as 1} \pt = \Spec Λ
\]
with $Λ = ℂ[λ]$ and $\deg λ = -1$.

\section{Conventions}

We consider all schemes as dg schemes.
In particular, all fiber products will be derived.
We will use the setup for ind-coherent sheaves on dg schemes described in \cite{Gaitsgory:2013:IndcoherentSheaves} and \cite{GaitsgoryRozenblyum:prelim:StudyInDAG}.

\section{Convolution}

Let $f\colon X → Y$ be a morphism of dg schemes.
We define the \emph{convolution product} on the category of ind-coherent sheaves on the groupoid $\Gpd{X}{Y} = X ×_Y X$ by
\[
    (\sheaf F, \sheaf G) \mapsto \sheaf F * \sheaf G \coloneq
    p_{13,*}\bigl(p_{12}^! \sheaf F \shriektensor p_{23}^! \sheaf G\bigr),
\]
where $p_{ij}\colon X ×_Y X ×_Y X → X ×_Y X$ are the projection maps.
Let $Δ\colon X → X×_YX$ be the diagonal morphism and $p_i\colon X×_YX → X$ the projection morphisms.

\begin{Lem}
    \label{lem:convolution_lemma}%
    Let $\sheaf F ∈ \catIndCoh{X}$ and $\sheaf G ∈ \catIndCoh{X×_YX}$.
    Then
    \[
        Δ_*\sheaf F * \sheaf G = p₁^!\sheaf F \shriektensor \sheaf G
    \]
    and
    \[
        \sheaf G * Δ_*\sheaf F = \sheaf G \shriektensor p₂^!\sheaf F.
    \]
\end{Lem}

\begin{proof}
    We will only prove the first statement.
    The second is completely analogous.
    Let $\tilde Δ\colon \Gpd{X}{Y} → \Gpd{X}{Y} ×_X \Gpd{X}{Y} = X ×_Y X ×_Y X$ be the diagonal.
    Thus we have the Cartesian square
    \[
        \begin{tikzcd}
            \Gpd{X}{Y} \arrow[r, "\tilde Δ"] \arrow[d, "p₁"] & \Gpd{X}{Y} ×_X \Gpd{X}{Y} \arrow[d, "p_{12}"] \\
            X \arrow[r, "Δ"] & \Gpd{X}{Y}
        \end{tikzcd}
    \]
    We obtain
    \begin{align*}
        Δ_*\sheaf F * \sheaf G
        & = p_{13,*}\bigl(p_{12}^! Δ_* \sheaf F \shriektensor p_{23}^! \sheaf G\bigr) & \\
        & = p_{13,*}\bigl(\tilde Δ_* p_{1}^! \sheaf F \shriektensor p_{23}^! \sheaf G\bigr) & \text{(base change)} \\
        & = p_{13,*}\tilde Δ_* \bigl(p_{1}^! \sheaf F \shriektensor \tilde Δ^!p_{23}^! \sheaf G\bigr) & \text{(projection formula)} \\
        & = \id_* \bigl(p_{1}^! \sheaf F \shriektensor \id^! \sheaf G\bigr) &  \\
        & = p_{1}^! \sheaf F \shriektensor \sheaf G. & \qedhere
    \end{align*}
\end{proof}

\begin{Thm}
    The category $\catIndCoh{X×_YX}$ is monoidal under convolution.
    The unit is $Δ_*ω_X$.
\end{Thm}

\begin{proof}
    We will only show that $Δ_*ω_X$ is a unit.
    Let $\sheaf G ∈ \catIndCoh{\Gpd{X}{Y}}$.
    Then by Lemma~\ref{lem:convolution_lemma},
    \[
        Δ_*ω_X * \sheaf G =
        p₁^! ω_X \shriektensor \sheaf G =
        ω_{\Gpd{X}{Y}} \shriektensor \sheaf G =
        \sheaf G,
    \]
    since $ω$ is a unit of $\shriektensor$.
\end{proof}

\begin{Cor}
    The functor $Δ_*$ is monoidal, i.e.~
    \[
        Δ_*(\sheaf F₁ \shriektensor \sheaf F₂) = Δ_* \sheaf F₁ * Δ_* \sheaf F₂, \qquad \sheaf F₁,\, \sheaf F₂ ∈ \catIndCoh{X}.
    \]
\end{Cor}

\begin{proof}
    Using Lemma~\ref{lem:convolution_lemma} and the projection formula,
    \[
        Δ_* \sheaf F₁ * Δ_* \sheaf F₂ =
        p₁^! \sheaf F₁ \shriektensor Δ_*\sheaf F₂ =
        Δ_*\bigl(Δ^!p₁^! \sheaf F₁ \shriektensor \sheaf F₂\bigr) =
        Δ_*\bigl(\sheaf F₁ \shriektensor \sheaf F₂\bigr).
        \qedhere
    \]
\end{proof}

\section{Koszul duality}

Let us now consider the special case of the space of based loops in $\as n$,
\[
    Ω₀\as n =
    \Gpd{\pt}{\as n} =
    \Spec k[λ₁,\dotsc, λ_n],
\]
where $\deg λ_i = -1$.
The dg scheme $Ω₀\as n$ has exactly one closed point given by $Δ\colon \pt → \Gpd{\pt}{\as n}$ corresponding to the quotient map $k[λ₁,\dotsc,λ_n] → k$.

We will need the following version of the Barr-Beck-Lurie Theorem, taken from \cite{Gunningham:thesis}.

\begin{Thm}
    \label{thm:BB-one-generator}%
    Let $\cat C$ be a cocomplete dg category with a single compact generator $M$ and set $R = \End_{\cat C}(M)^{\mathrm{op}}$.
    Then the assignment $N \mapsto \Hom_{\cat C}(M, N)$ defines an equivalence $\cat C \cong \catModules{R}$.
\end{Thm}

We note that the category $\catIndCoh{\Gpd{\pt}{\as n}}$ is compactly generated by $Δ_*ω_{\pt} = Δ_*k_{\pt}$.

\begin{Thm}
    \label{thm:Koszul-computation}%
    There is an isomorphism of (dg) algebras
    \[
        \End_{\catIndCoh{\Gpd{\pt}{\as n}}}(Δ_*ω_{\pt}) \cong k[u₁, \dotsc, u_n], \qquad \deg u_i = 2.
    \]
\end{Thm}

\begin{proof}
    See \cite{MO:ComputingMultiplicationInExtKoszul}.
    The fact that this computes the right algebra structure is contained in the proof of \cite[Theorem~1.7]{SuarezAlvarez:2004:HiltonHeckmannArgument}.
\end{proof}

Thus we recover \cite[Corollary 5.1.10]{ArinkinGaitsgory:2015:SingularSupport}:

\begin{Cor}[Koszul duality for exterior algebra]
    \label{thm:Kozul-duality}%
    There is a canonical equivalence of dg categories
    \[
        \catIndCoh{\Gpd{\pt}{\as n}} \isoto \catModules{k[u₁,\dotsc,u_n]},
    \]
    sending $Δ_*ω_{\pt}$ to $k[u₁,\dotsc, u_n]$ and $\O_{\Gpd{\pt}{\as n}}$ to $k$[shift?].
    This equivalence restricts to an equivalence
    \[
        \catModulesFG{k[λ₁,\dotsc, λ_n]} = \catCoh{\Gpd{\pt}{\as n}} \isoto \catModulesFG{k[u₁,\dotsc,u_n]}
    \]
    and
    \[
        \catModules{k[λ₁,\dotsc, λ_n]} = \catQCoh{\Gpd{\pt}{\as n}} \isoto \catModules[][\{0\}]{k[u₁,\dotsc,u_n]},
    \]
    where the latter category denotes the full subcategory of modules whose cohomological (set-theoretic) support is contained in $\{0\} \subset \as n = \Spec k[u₁,\dotsc,x_n]$.
\end{Cor}

\begin{proof}
    The first statement follows directly from Theorems~\ref{thm:BB-one-generator} and~\ref{thm:Koszul-computation}.
    The second statement is just the restriction to compact objects.
    The generator $k[λ₁,\dotsc,λ_n]$ is sent to a shift of $k$ under Koszul duality.
    We note that $k$ has support $\{0\}$ as a $k[u₁,\dotsc,u_n]$ module, which implies the third equivalence.
\end{proof}

\begin{Rem}
    The dg scheme $\Gpd{\pt}{\as n}$ is quasi-smooth, but not smooth.
    Hence perfect sheaves and coherent sheaves do not coincide on it, and thus neither do their ind-completions $\catQCoh{\Gpd{\pt}{\as n}}$ and $\catIndCoh{\Gpd{\pt}{\as n}}$.
    The fully faithful functor
    \[
        Ξ\colon \catQCoh{\Gpd{\pt}{\as n}} \hookrightarrow \catIndCoh{\Gpd{\pt}{\as n}}
    \]
    of \cite{Gaitsgory:2013:IndcoherentSheaves} corresponds under Koszul duality to the inclusion
    \[
        \catModules[][\{0\}]{k[u₁,\dotsc,u_n]} \hookrightarrow \catModules{k[u₁,\dotsc,u_n]}.
    \]
    For a more in-depth discussion of we refer to the lecture \cite{conference:2014:GLJerusalem:Arinkin:QASingSuppKoszul}.
\end{Rem}

\begin{Rem}
    Theorem~\ref{thm:BB-one-generator} should be upgradeable to an equivalence of symmetric monoidal categories (under some assumptions).
    Thus Corollary~\ref{thm:Kozul-duality} can be upgraded to an equivalence of symmetric monoidal categories.
\end{Rem}


\section{Hochschild cohomology}

\printbibliography
\end{document}
