%type: notes
%title: Spectral Eisenstein series functors
%tags: representation theory, geometric Langlands, local systems
\documentclass[english]{short-notes}

\usepackage{math-alg,math-ag, math-gl}
\usepackage{xparse}

\addbibresource{global.bib}
%\bibliography{global.bib}


\title{spectral eisenstein series functors}
\author{Clemens Koppensteiner}

\begin{document}

\newcommand\sym{\mathrm{sym}}
\newcommand\STs[2]{\operatorname{ST}(#2,#1)}
\newcommand\STf[2]{\operatorname{ST}_{#2,#1}}

\newcommand\cusp{\mathrm{cusp}}
\newcommand\irred{\mathrm{irred}}
\newcommand\red{\mathrm{red}}
\newcommand\globNilp{\mathcal{N}}
\newcommand\twist[1]{\operatorname{T}_{#1}}

\maketitle

\tableofcontents
\bigskip

The following is mainly taken from \cite[Section~12]{ArinkinGaitsgory:arXiv:v2:SingularSupport}.

Let $C$ be a curve over a $ℂ$, $G$ a reductive group.
We are interested in the dg category $\catIndCoh{\Loc_G}$ of ind-coherent sheaves on the derived moduli stack of local $G$-systems on $C$.

Let $P$ be a parabolic subgroup of $G$ and $L$ the corresponding Levi.
We write
\[
    p_P\colon \Loc_P → \Loc_G, \qquad q_P\colon \Loc_P → \Loc_L.
\]
The map $p_P$ is schematic and proper and the map $q_P$ is quasi-smooth \cite[Lemma~12.2.2]{ArinkinGaitsgory:arXiv:v2:SingularSupport}.

\begin{Q}
    Does $q_P$ have more properties than just being quasi-smooth?
    What is its relative dualizing sheaf?
    
    Maybe making the proof of \cite[Lemma~12.2.2]{ArinkinGaitsgory:arXiv:v2:SingularSupport} explicit leads to nicer properties of $q$?
    See also \cite[Proposition~9.4.5]{ArinkinGaitsgory:arXiv:v2:SingularSupport}.
\end{Q}

\begin{Def}
    The \emph{spectral Eisenstein series functor} for $P$ is
    \[
        \Eis_P^G = (p_P)_*^\IndCoh ∘ q_P^!\colon \catIndCoh{\Loc_M} → \catIndCoh{\Loc_G}.
    \]
\end{Def}

The functor $(p_P)_*^\IndCoh$ is defined in~\cite[Proposition~3.1.1]{Gaitsgory:preprint:IndcoherentSheaves} and the functor $q_P^!$ exists by \cite[Theorem~5.2.2]{Gaitsgory:preprint:IndcoherentSheaves}.

When the group $G$ is clear from context, we set $\Eis_P = \Eis_P^G$.
If there is only on parabolic $P$, then $p = p_P$, etc.

\begin{Q}
    We are interested in the following questions:
    \begin{itemize}
        \item 
            Let $P$ and $Q$ be two parabolics with equivalents Levis.
            Are the functors $\Eis_P$ and $\Eis_Q$ isomorphic?
            (Compare Proposition~3.1.9 in Sam's thesis.)
        \item 
            Are the images of the $\Eis$ for non-equivalent Levis orthogonal?
    \end{itemize}
\end{Q}

\section{adjoints}

\begin{Q}
    Does the functor $\Eis$ have adjoints?
\end{Q}

\begin{Rem}\leavevmode
    \begin{itemize}
        \item $p_*^\IndCoh$ has the right adjoint $p^!$ \cite[3.3.7]{Gaitsgory:preprint:IndcoherentSheaves}.
        \item There is an adjunction $(q^{\IndCoh,*},q_*^\IndCoh)$ \cite[Corollary~1.2.5]{ArinkinGaitsgory:arXiv:v2:SingularSupport}.
        \item For general maps, $q^!$ does not need to be adjoint to $q_*^\IndCoh$.
    \end{itemize}
\end{Rem}

\begin{Prop}[{\cite[Corollary~12.2.3]{ArinkinGaitsgory:arXiv:v2:SingularSupport}}]
    The functor $\Eis_P$ sends $\catCoh{\Loc_M}$ to $\catCoh{\Loc_G}$.
\end{Prop}

\begin{Cor}
    $\Eis_P^G$ has a (continuous) right adjoint.
    We will denote it by $\CT_P^G = \CT_P$ (for \emph{spectral constant term functor}).
\end{Cor}

\begin{proof}
    The functor $\Eis$ is continuous, so by the adjoint functor theorem \cite[Corollay~5.5.2.9]{Lurie:2009:HigherToposTheory} it has a right adjoint.
    By \cite[Proposition~5.5.7.2]{Lurie:2009:HigherToposTheory}, the right adjoint is continuous.
\end{proof}

\begin{Q}
    Is $\CT$ also a left adjoint?
\end{Q}

Let $\sheaf K_P^\dual$ denote the dual of $q_P^{\mathrm{QCoh},!}(\O_{\Loc_L})$ (in $\catQCoh{\Loc_P}$).

\begin{Lem}
    The functor $\CT_P$ is equal to
    \[
        \sheaf F \mapsto q_*^\IndCoh(\sheaf K_P^\dual \otimes p^!(\sheaf F)).
    \]
\end{Lem}

\begin{Rem}%
    \label{rem:symmetric}%
    Assume there exists $\sheaf L ∈ \catIndCoh{\Loc_P}$ such that $\sheaf K = \sheaf L \otimes \sheaf L$ (such $\sheaf L$ does not need to exist in general; for example if $\sheaf K$ is in odd degree).
    Then setting $\Eis(\sheaf F) = p_*^\IndCoh(\sheaf L^\dual \otimes q^!(\sheaf F))$ yields the much more symmetric $\CT(\sheaf G) = q_*^\IndCoh(\sheaf L^\dual \otimes p^!(\sheaf G))$.
    It would also ensure that $\Eis ∘ {\mathbb D} = \mathbb D ∘ \Eis$, at least on $\cat{Coh}$ (see~\cite[Corollary~9.5.9]{Gaitsgory:preprint:IndcoherentSheaves}).

    Of course, it would be even better if $\sheaf K = \sheaf L = \O_{\Loc_P}$.
\end{Rem}

\begin{Notation}%
    There are several different ways to define the Eisenstein series functor (and consequently the constant term functor).
    We will use the following notation:
    \begin{itemize}
        \item $\Eis = \Eis^! = p_*^\IndCoh ∘ q^!$ with $\CT = \CT^! = q_*^\IndCoh(\sheaf K_P^\dual \otimes p^!({-}))$.
            This variant is used in \cite{ArinkinGaitsgory:arXiv:v2:SingularSupport}.
        \item $\Eis^* = p_*^\IndCoh ∘ q^*$ with $\CT^* = q_*^\IndCoh ∘ p^!$.
            This variant is used in \cite{Gaitsgory:preprint:OutlineOfGLCForGL2} and denoted simply $\Eis_{\mathrm{spec}}$ there.
        \item $\Eis^\sym$ and $\CT^\sym$ for the versions defined in Remark~\ref{rem:symmetric} (if existing).
    \end{itemize}
    Analogous notation is applied for functors derived from these.
\end{Notation}

\section{the \enquote{steinberg functor}}

In this section let $P$ and $Q$ be two parabolic subgroups of $G$ with Levis $L$ and $M$ respectively.

\begin{Def}
    The \emph{spectral Steinberg functor} is
    \[
        \STf P Q = \CT_Q ∘ \Eis_P \colon \catIndCoh{\Loc_L} → \catIndCoh{\Loc_M}.
    \]
    The \emph{spectral Steinberg stack} is
    \[ 
        \STs P Q =
        \Loc_P ×_{\Loc G} \Loc_Q.
    \]
    The two projections from $\STs P Q$ will be denoted by $π_P$ and $π_Q$.
    Further, let $α_P = q_P ∘ π_P$ and analogously $α_Q$.
\end{Def}

\begin{Cor}
    \[
        \STf P Q ({-}) = 
        (α_Q)_*^\IndCoh \bigl( π_Q^*(\sheaf K_Q^\dual) \otimes α_P^!({-})\bigr).
    \]
\end{Cor}

\begin{proof}
    \begin{align*}
        \CT_Q ∘ \Eis_P (\sheaf F) &=
        (q_Q)_*^\IndCoh\biggl( \sheaf K_Q^\dual \otimes \bigl( p_Q^! ∘ (p_P)_*^\IndCoh ∘ q_P^!(\sheaf F)\bigr)\biggr) \\ & =
        (q_Q)_*^\IndCoh\biggl( \sheaf K_Q^\dual \otimes \bigl( (π_Q)_*^\IndCoh ∘ π_P^! ∘ q_P^!(\sheaf F)\bigr)\biggr) \\ & =
        (α_Q)_*^\IndCoh \bigl( π_Q^*(\sheaf K_Q^\dual) \otimes α_P^!(\sheaf F)\bigr),
    \end{align*}
    where the first equality is (proper) base change and the second one the compatibility of the pushforward with the action of $\cat{QCoh}$.
\end{proof}

\begin{Rem}
    In the symmetric situation of Remark~\ref{rem:symmetric}, one gets
    \[
        \STf P Q ^\sym({-}) =
        \CT^\sym_Q ∘ \Eis^\sym_P ({-}) = 
        (α_Q)_*^\IndCoh \biggl( \bigl( π_Q^*(\sheaf L_Q^\dual) \otimes π_P^*(\sheaf L_P^\dual) \bigr) \otimes α_P^!({-})\biggr),
    \]
    while
    \[
        \STf P Q ^*({-}) =
        \CT^*_Q ∘ \Eis^*_P ({-}) = 
        (α_Q)_*^\IndCoh \bigl( π_P^*(\sheaf K_P) \otimes α_P^!({-})\bigr).
    \]
\end{Rem}

\begin{Q}
    Is
    \[
        \sheaf K_Q^\dual =
        \mathbb D_{\Loc_Q}^{\mathrm{naive}} q_Q^{\mathrm{QCoh},!}(\O_{\Loc_M}) =
        q_Q^* (\mathbb D_{\Loc_M}^{\mathrm{naive}}\O_{\Loc_M}),
    \]
    and hence
    \[
        π_Q^*(\sheaf K_Q^\dual) =
        α_Q^*(\mathbb D_{\Loc_M}^{\mathrm{naive}}\O_{\Loc_M})?
    \]
    Further, is $\mathbb D_{\Loc_M}^{\mathrm{naive}}\O_{\Loc_M} = \O_{\Loc_M}$?
\end{Q}

\begin{Q}
    Does $\STs P Q$ have a natural (equidimensional) stratification?
\end{Q}

\subsection{a \texorpdfstring{$G$}{G}-equivariant formulation}

Let $[P] \cong G/P$ denote the partial flag variety corresponding to the parabolic $P$, i.e.~the set of all conjugates of $P$.

We consider the \enquote{Betti version} of the moduli of local systems.
Let $\Loc_G'$ be the non-stacky moduli.
If $X$ is an elliptic curve this is the (derived) fiber product $\{1\} ×_G G²$ (with $[{,}]\colon G² → G$).
By definition, $\Loc_G = \rquot{\Loc_G'}G$.

\begin{Def}
    Let $\Loc_{[P]}'$ be the dg scheme over $[P]$ with fibers $\Loc_{P'}'$ over $P' ∈ [P]$.
\end{Def}

\begin{Q}
    How to define/construct this in a formally correct way?
\end{Q}

\begin{Claim}
    There is a canonical isomorphism of derived stacks $\Loc_P \cong \rquot{\smash{\Loc_{[P]}'}}G$.
\end{Claim}

We set 
\[
    \Loc_{[P∩Q]}' = \Loc_{[P]}' ×_{\Loc_G'} \Loc_{[Q]}'.
\]
Informally, this is the dg scheme over $[P] × [Q]$ with fibers $\Loc_{P'∩Q'}$ over $(P',Q') ∈ [P]×[Q]$.

\begin{Claim}
    There is a canonical isomorphism of derived stacks $\STs P Q \cong \rquot{\smash{\Loc_{[P∩Q]}'}}G$.
\end{Claim}

Note that the $G$-orbits on $[P]×[Q]$ are indexed by ${}_QW_P = \dquot{W_Q}{W}{W_P}$.
Thus orbits on $[P]×[Q]$ induce a stratification of $\Loc_{[P∩Q]}'$ indexed by $_QW_P$, and hence one of $\STs P Q$.
The strata are denoted $\STs P Q _w$

\begin{Conjecture}
    The strata are equidimensional.
\end{Conjecture}

\begin{Q}
    If $({}^wP,Q) ∈ [P]×[Q]$ is a point in the orbit indexed by $w$, is $\STs PQ = \Loc_{^wP∩Q}$?
\end{Q}

\begin{Q}
    Can the stratification of $\STs P Q$ be obtained without the detour through $\Loc'$?
    How to obtain a stratification for the \enquote{de Rham moduli stack}?
\end{Q}

Next we should define functors ${\STf P Q} _w\colon \catIndCoh{\Loc_L} → \catIndCoh{\Loc_M}$ given by restriction of $α$ and $β$ to $\STs P Q _w$.

\begin{Q}
    Do the functors ${\STf P Q}_w$ respect nilpotent singular support?    
\end{Q}

\begin{Q}
    Is $\STf P Q = \bigoplus_w {\STf P Q}_w$?
\end{Q}

\subsection{a general filtration}

A choice of map $\{\mathrm{pt}\} → C$ yields a morphism $\Loc_G(C) → \mathrm BG$ via pullback (and the forgetful map $\Loc_G → \Bun_G$.
Thus there is a morphism
\[
    γ\colon \STs PQ → \mathrm BQ ×_{\mathrm BG} \mathrm BP = {}_QW_P.
\]
\begin{Def}
    For $w ∈ {}_QW_P$ set
    \begin{align*}
        \STs PQ_w & = γ^{-1}(w), \\
        \STs PQ_{\overline  w} &= γ^{-1}(\overline{\{w\}}).
    \end{align*}
\end{Def}
The set $_QW_P$ is partially ordered by $w₁ \le w₂$ if $\overline{\{w₁\}} ⊆ \overline{\{w₂\}}$.
Hence the stack $\STs PQ$ is filtered by the closed subspaces $\STs PQ_{\overline w}$.
This in turn induces a filtration on $\STf PQ$.

\begin{Q}
    What can we say about this filtration?
\end{Q}

\section{outline}

% try to get some more uniform notation
\RenewDocumentCommand\Eis{omo}{\operatorname{Eis}_{#2}^{\IfValueTF{#3}{#3\IfValueT{#1}{,#1}}{\IfValueT{#1}{#1}}}}
\RenewDocumentCommand\CT{omo}{\operatorname{CT}^{#2\IfNoValueF{#1}{,#1}}_{\IfNoValueF{#3}{#3}}}
\RenewDocumentCommand\STf{omm}{\operatorname{ST}_{#2\IfNoValueF{#1}{,#1}}^{#3}}
\renewcommand\STs[3][]{\operatorname{ST}(#3,#2)_{#1}}
\newcommand\STm{\operatorname{ST}}
\newcommand\relW[2]{W_{#1}^{#2}}
\newcommand\DSerre[1][]{\mathbf{D}^{\mathrm{Serre}}_{#1}}
\newcommand\op{\mathrm{op}}

\subsection{the problem}

By \cite[Corollary~12.3.10]{ArinkinGaitsgory:arXiv:v2:SingularSupport}, the essential images of of $\catQCoh{\Loc_L} ⊂ \cat{IndCoh}_{\mathrm{Nilp}}(\Loc_L)$ under the functor $\Eis P[G]$ for all parabolic subgroups $P$ of $G$ (including $P=G$) generate $\cat{IndCoh}_{\mathrm{Nilp}}(\Loc_G)$.

The goal is to make this statement more precise:
\begin{itemize}
    \item Does it suffice to look at the images of the \enquote{cuspidal} part $\catQCoh{\Loc_L^{\mathrm{irred}}}$?
    \item For two equivalent Levis, are the images the same?
    \item Are the images (of the cuspidal parts) orthogonal?
    \item Can the images (of the cuspidal parts) be described as modules of a monad on $\catQCoh{\Loc_L}$?
\end{itemize}

A further question: How important is the nilpotent singular support condition for this decomposition?

\subsection{the functors}

To show the orthogonality of the images, we have to show that $\Hom(\Eis P \sheaf F,\, \Eis Q \sheaf G) \ne 0$ for cuspidal $\sheaf F$, $\sheaf G$ only if $M$ and $N$ are equivalent Levis.
To do so we show that $\CT P ∘ \Eis Q = 0$ if $M$ and $N$ are not equivalent.
Here $\CT P$ is a two-sided adjoint to $\Eis P$.

This raises the question whether a two-sided adjoint even exists.
By general abstract nonsense, $\Eis P$ has a right adjoint, which is easy to describe.
The question of a left adjoint (which does not necessarily need to equal the right adjoint for our purposes) is harder.

In any case, we need to study the \emph{Steinberg functor} $\STf P Q = \CT Q[G] ∘ \Eis P[G]$.

\subsubsection{Adjoints}

\begin{Claim}
    \label{claim:CT-formula}
    The right adjoint to $\Eis P[G]$ is given by
    \[
        \CT P[G]\colon \catIndCoh{\Loc_G} → \catIndCoh{\Loc_L}\colon \sheaf F \mapsto ω_{\Loc_L} \otimes q_*^\IndCoh p^!(\sheaf F).
    \]
\end{Claim}

The endofunctor $ω_{\Loc_L} \otimes {-}$ of $\catIndCoh{\Loc_L}$ will be denoted $\twist P$.

Unfortunately, there is no reason to think that the functor $\CT P[G]$ does not need to send coherent sheaves to coherent sheaves.

\begin{Q}
    Does $\CT P$ preserve compactness for $X = \ps 1$ or $X$ an elliptic curve?
\end{Q}

\begin{Lem}
    There is an isomorphism of functors $\Eis[!]P ∘ \DSerre[\Loc_L] \cong \DSerre[\Loc_G] ∘ \Eis[*,\op]P$ as functors $\catIndCoh{\Loc_L}^\dual → \catIndCoh{\Loc_G}$ (using the notation of \cite[Section~2.3.2]{Gaitsgory:preprint:GL.DGcat}).
\end{Lem}

\begin{proof}
    By \cite[Section~9.2.3]{Gaitsgory:preprint:IndcoherentSheaves},
    \[
        \Eis[!]P ∘ \DSerre \cong
        {(p^!)}^\dual ∘ {(q_*^\IndCoh)}^\dual ∘ \DSerre =
        \CT[*,\dual]P ∘ \DSerre.
    \]
    By \cite[Lemma~2.3.3]{Gaitsgory:preprint:GL.DGcat} the right adjoint $\CT[*]P$ of $\Eis[*]P$ is isomorphic to $\Eis[*,\op,\dual]P$.
    Hence $\CT[*,\dual]P \cong \Eis[*,\op]P$.
\end{proof}

\subsubsection{Cuspidal sheaves}

\begin{Def}
    A sheaf $\sheaf F ∈ \catIndCoh{\Loc_G}$ is \emph{cuspidal} if it is in the kernel of $\CT P$ (or equivalently $\CT[*]P$) for all for all proper parabolic subgroups $P$ of $G$.
    The full subcategory of cuspidal sheafs will be denoted $\catIndCoh{\Loc_G}_\cusp$.
\end{Def}

By definition (compare~\cite[Remark~5.2.8.7]{Lurie:2009:HigherToposTheory}) this means that $\catIndCoh{\Loc_G}_\cusp$ is right orthogonal to the subcategory of $\catIndCoh{\Loc_G}$ that is generated by the essential images of all $\Eis P$ for proper parabolics $P$.
Essentially, we want to show that it is also left orthogonal.

By \cite[(12.3) and Proposition~12.3.3]{ArinkinGaitsgory:arXiv:v2:SingularSupport} we have
\begin{equation}%
    \label{eq:QCohIrredIdentification}
    \catIndCoh[\globNilp]{\Loc_G}_\cusp \cong \catIndCoh[\globNilp]{\Loc_G^\irred} \cong \catQCoh{\Loc_G^\irred}.
\end{equation}

\begin{Conjecture}
    There exists a decomposition
    \[
        \catIndCoh[\globNilp]{\Loc_G} = \bigoplus_L \Eis{P(L)}\bigl(\catQCoh{\Loc_L^\irred}\bigr),
    \]
    where $L$ ranges over the conjugacy classes of Levis of $G$ (including $L=G$), $P(L)$ is a parabolic subgroup representing $L$ and $\catQCoh{\Loc_L^\irred}$ is identified with a subcategory of $\catIndCoh[\globNilp]{\Loc_L}$ via \eqref{eq:QCohIrredIdentification}.
\end{Conjecture}

\subsection{a filtration of the steinberg functor}

The functor $\STf P Q $ can be described via pull-push over the derived stack $\STs P Q$.
Concretely we have (assuming Claim~\ref{claim:CT-formula})
\[
    \STf PQ = \twist Q ∘ (α_Q)_*^\IndCoh ∘ α_P^! \sheaf F.
\]
By definition we have a map $π\colon \STs PQ → \relW PQ$ which induces a filtration on $\STf PQ$.
For $w ∈ \relW PQ$ let $\STs[w]PQ = π^{-1}(w)$.
Note that this is a locally closed substack.
Let $j_w\colon \STs[w]PQ \hookrightarrow \STs PQ$ denote the inclusion and let $α_{P,w} = α_P ∘ j_w$ (resp.~$α_{Q,w} = α_Q ∘ j_w$) be the restriction of $α_P$ (resp.~$α_Q$).
Define
\[
    \STf[w] PQ =
    \twist Q ∘ {(α_{Q,w})}_*^\IndCoh ∘ α_{P,w}^! =
    \twist Q ∘ {(α_Q)}_*^\IndCoh ∘ {(j_w)}_*^\IndCoh ∘ j_w^! ∘ α_P^! \sheaf F.
\]

\begin{Claim}\label{claim:ST_as_iterated_extension}
    $\STf PQ(\sheaf F)$ is an iterated extension of the sheaves $\STf[w] PQ(\sheaf F)$.
\end{Claim}

Thus we need to describe the functors $\STf[w]PQ$.

\subsection{a description of the subquotient functors}

\begin{Claim}
    \[
        \STs[w]PQ = \Loc_{^wP∩Q}.
    \]
\end{Claim}

We describe $\STf[w]PQ$ as $\Eis R[M] ∘ \CT R[^wL]$ over a common Levi of $^wL$ and $M$.
If $L$ and $M$ are not equivalent (and hence the common Levi proper) this functor has to vanish on \emph{cuspidal} sheaves by definition.
More precisely, we need to show the following claim.

\begin{Claim}
    If $\Eis{P_M}[M] ∘ \CT{Q_L}[L] \sheaf F = 0$, then $\STf[w] PQ \sheaf F = 0$ (where $w$ corresponds to the orbit of $(P,Q)$).
\end{Claim}

\begin{Cor}
    If $^wL$ is not a Levi of $M$, then $\STf[w] PQ = 0$ on cuspidal sheaves.
\end{Cor}

\begin{Cor}
    If $L$ is not conjugate to a Levi of $M$, then $\STf PQ = 0$ on cuspidal sheaves.
\end{Cor}

\begin{proof}
    In this case all $\STf[w] PQ$ vanish and hence $\STf PQ = 0$ by Claim~\ref{claim:ST_as_iterated_extension}.
\end{proof}

\subsection{describing the images}

For $L$ a Levi of $G$, we write $\catIndCoh[\globNilp]{\Loc_G}^{[L]}$ for the full subcategory generated by the image of $\catIndCoh[\globNilp]{\Loc_L}_\cusp$ under $\Eis P[G]$.
We have adjoint functors
\[
    \Eis P[G] \colon \catIndCoh[\globNilp]{\Loc_L}_\cusp \leftrightarrows \catIndCoh[\globNilp]{\Loc_G}^{[L]} \cocolon \CT P[G]
\]
and hence a monad $\STm = \STf PP = \CT P ∘ \Eis P$ on $\catIndCoh[\globNilp]{\Loc_L}_\cusp$.

\begin{Cor}
    There is an equivalence of categories
    \[
        \catIndCoh[\globNilp]{\Loc_G}^{[L]} \cong \catIndCoh[\globNilp]{\Loc_L}_\cusp^{\STm}.
    \]
\end{Cor}

\begin{proof}
    The functor $\CT P[G]$ restricted to $\catIndCoh[\globNilp]{\Loc_G}^{[L]}$ is (tautologically) conservative.
    Hence the result follows from the Bar--Beck--Lurie theorem \cite[Theorem~6.2.0.6]{Lurie:2012-draft:HigherAlgebra}.
\end{proof}

\begin{Q}
    Can the monad $\STm$ be described in a more direct way, \enquote{internal} to $\catIndCoh[\globNilp]{\Loc_L}_\cusp$?
\end{Q}

\section{describing cuspidal sheaves}

\begin{Q}
    Is there a nice description or decomposition of the category $\catIndCoh[\globNilp]{\Loc_G}_{\cusp}$?
    Maybe there is a natural decomposition of the space $\Loc_G^\irred$?
    If not in general, then maybe over $\ps 1$ or an elliptic curve?
\end{Q}

\section{a filtration}

The global nilpotent cone $\globNilp$ has a natural stratification by orbits.

\begin{Q}
    Can one describe the induced filtration of $\catIndCoh[\globNilp]{\Loc_G}$?
\end{Q}

\printbibliography
\end{document}
